\documentclass[colorlinks,11pt]{moderncv}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[scaled]{helvet}
\renewcommand\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

\usepackage{url}

\moderncvstyle{classic}
\moderncvcolor{blue}


\name{Thomas Ekstr{\" o}m}{Hansen}
\title{Final-year MSci student}
\address{14 Chambers Place}{St Andrews}{KY16 8RQ}

\recipient{Dr Chris Jefferson}{Director of Postgraduate Research\\School of
           Computer Science\\$\pi$ Potato Avenue\\St Andrews}
\date{}
\opening{Dear Dr Jefferson,}
\closing{Yours sincerely,}

\begin{document}
    
    \vspace*{-1cm}
    
    \makelettertitle
    
    %Please state briefly your reasons for applying for postgraduate study at St Andrews and what in your educational background makes you a particularly suitable candidate for graduate study in your chosen programme
    
    I am applying to St Andrews for a PhD in Computer Science. Being in the
    final year of my Integrated Masters degree here, St Andrews is a place I
    have grown very familiar with and fond of. The School of Computer Science is
    one of the leading research institutions in Europe and I have had a great 5
    years here during my undergraduate degree. In addition to the school being a
    friendly and welcoming work environment, it also houses great staff. Dr
    Brady is one of the leading experts in type systems and functional
    programming languages. His programming language \textsc{Idris} is, as far as
    I know, the current state-of-the-art in useable dependently typed
    programming languages. Therefore, given my own interest in type systems and
    functional programming, I believe he would work very well as a supervisor
    for my PhD proposal concerning dependent types and formal methods.
    Furthermore, I know the school has an interest in formal methods and that Dr
    Bowles is interested in the modelling and verification of critical systems
    as well as correctness proofs. Given these two research groups and
    interested academics, I believe St Andrews would be an excellent place to
    undertake my proposed PhD project.
    
    I would like to bring your attention to a couple of details in my
    educational background which I believe make me particularly suitable for
    this research position. During my undergraduate at the university, I took
    modules in both Logic and Software Verification, and Critical Systems
    Engineering. I did well in both, achieving a 17.9 and a 19.6 in them
    respectively. My Major Software Project dissertation in my
    4\textsuperscript{th} year made use of \textsc{Idris} to construct provably
    correct certificates for embedded systems' software. This also went well, I
    received a 17 for it. Additionally, I have been on the Dean's List for the
    past 3 years. Finally, I frequently talk and interact with many of the
    current staff and PhD students, and I am very interested in the teaching and
    tutoring aspects in which the school involves its PhD students. Therefore, I
    believe I am highly qualified to undertake my proposed PhD and that I would
    fit in well in the department.
    
    I look forward to discussing my application with you.
    
    \makeletterclosing

\end{document}
